Investigating the impacts of COVID-19 on team wins and social media engagement in the EPL and Manchester United

### 2022-04-27 outputs/paper/paper.Rmd
This is the final commit to make sure all files are up to date in order for the paper to be fully functional.

\tableofcontents

@@

# Introduction
The English Premier League is renowned to be one of the most watched soccer leagues in the world. To date, the league is broadcasted in over 200 countries which spans over 4 billion people (@citeRichards). With such a large market, teams in the EPL are continually able to attract some of the best players in the world. This is pivotal because, not only will the league be attracting more engagement, but the level of competition will be more intense. At the start of 2020, the COVID-19 pandemic had halted a lot of major activities around the world. With the virus spreading quick, the organizers of the EPL halted the premier league football between March to June. After the initial suspension of the games, teams were allowed to play again but behind closed doors meaning no attendance and stricter protocols. We chanced upon an article on ESPN, that essentially highlights the loss of home advantage arising from games with no fans. Pivotally, they pointed out that Liverpool, who were the 2019-2020 season winners, had lost 6 games at home during the COVID era. Prior to that, they were on a 68 games win streak at home (@citeHamilton). We use @citetidyverse, @citegridExtra, @citehrbrthemes, @citefitdistrplus, @citereadxl and @citerms to prove how Covid has affected teams in the EPL both on and off the field.With the help of @citeknitr, this paper was made possible to print in a pdf format.

The paper looks to compare the effect fans had when describing team's win and as the statistic proves that there is a difference. We can see clearly from the table how Liverpool were champions in 2019-2020 season with 99 points while in the 2020-2021 season, Liverpool finished in 3rd place with 69 points. That shows a 30 point gaps. Although we are not looking at Liverpool, they are a perfect example of how Covid has affected their performances without fans. In regards to how Covid has affected Twitter Engagement for Manchester United, the highest total engagement by month has increased by 90% from a number of roughly 5 million to 9.5 million. Although twitter engagement has nothing to do with the team's performance, it is still interesting how fans are finding different ways to support their team during tough times. As majority of the people were working from home, the high engagement is justified.

@@

# Exploratory Data Analysis
This investigation will consist of 5 data sets and focus on teams in Premier League (in England). To assess the first question, we will be working with data sets 1-2. For reference, the 2019/2020 season will be associated with Pre-Covid period and 2020/2021 season will be associated with COVID period and the average season has 38 games.

@@

In order to run a logistic regression, the response variable needs to be of value 0 <y<1. We need to obtain a new response variable consisting of Wins. To do so, we will create new variable: Prop, which is the number of wins over matches played. Given that a team can't possibly win greater than 38 games, we are assured that our response will be less than 1. Another key change was to assign attendance according to: "High", "Medium" and "Low" which we have assigned according to the upper, median, and lower quantile of attendance variable. Same can be said for the 2020/2021, which we will assign "Very few" and "Some" which by its name, represents teams with very few fans and teams with some fans.

@@

Prop here represents the proportion of team wins with respect to total matches played by the team. 
GF and GA represents the goals scored and conceded by the team
Attendance here is a categorical variable that assigning, "High","Medium", "Low" according to the 25%,50% and 75% of variable attendance. Likewise is applied for the 2020/2021 season, where the only difference is that, "Some" and "Very Few" is applied to teams playing with some fans and teams playing very few fans in the stadium.

>>>>

\newpage
\tableofcontents
\newpage

@@

# Introduction
The English Premier League is renowned to be one of the most watched soccer leagues in the world. To date, the league is broadcasted in over 200 countries which spans over 4 billion people (@citeRichards). With such a large market, teams in the EPL are continually able to attract some of the best players in the world. This is pivotal because, not only will the league be attracting more engagement, but the level of competition will be more intense. At the start of 2020, the COVID-19 pandemic had halted a lot of major activities around the world. With the virus spreading quick, the organizers of the EPL halted the premier league football from March, 2020 to June, 2020. After the initial suspension of the games, teams were allowed to play again but behind closed doors meaning no attendance and stricter protocols. We chanced upon an article on ESPN, that essentially highlights the loss of home advantage arising from games with no fans. Pivotally, they pointed out that Liverpool, who were the 2019-2020 season winners, had lost 6 games at home during the COVID era. Prior to that, they were on a 68 games win streak at home (@citeHamilton). We use @citetidyverse, @citegridExtra, @citehrbrthemes, @citefitdistrplus, @citereadxl and @citerms to prove how Covid has affected teams in the EPL both on and off the field. With the help of @citeknitr, this paper was made possible to print in a pdf format.

The paper looks to compare the effect fans had when describing team's win and as the statistic proves that there is a difference. We can see clearly from the table how Liverpool were champions in 2019-2020 season with 99 points while in the 2020-2021 season, Liverpool finished in 3rd place with 69 points. That shows a 30 point gap. Although we are not looking at Liverpool, they are a perfect example of how Covid has affected their performances without fans. In regards to how Covid has affected Twitter Engagement for Manchester United, the highest total engagement by month has increased by 90% from a number of roughly 5 million to 9.5 million. Although twitter engagement has nothing to do with the team's performance, it is still interesting to notice how fans are finding different ways to support their teams during tough times. As majority of the people were working from home, the high engagement is justified.

@@

# Exploratory Data Analysis
This investigation will consist of 6 data sets and focus on teams in Premier League (in England). To assess the first question, we will be working with data sets 1-2. For reference, the 2019/2020 season will be associated with Pre-Covid period and 2020/2021 season will be associated with COVID period and the average league season has 38 games.

@@

In order to run a logistic regression, the response variable needs to be of value 0 < y < 1. We need to obtain a new response variable consisting of Wins. To do so, we will create new variable: Prop, which is the number of wins over matches played. Given that a team can't possibly win greater than 38 games, we are assured that our response will be less than 1. Another key change was to assign attendance according to: "High", "Medium" and "Low" which we have assigned according to the upper, median, and lower quantile of attendance variable. Same can be said for the 2020/2021, which we will assign "Very few" and "Some" which by its name, represents teams with very few fans and teams with some fans.

@@

Prop here represents the proportion of team wins with respect to total matches played by the team. 
GF and GA represents the goals scored and conceded by the team
Attendance here is a categorical variable that assigning, "High","Medium", "Low" according to the 25%,50% and 75% of variable attendance. Likewise is applied for the 2020/2021 season, where the only difference is that, "Some" and "Very Few" is applied to teams playing with some fans (more than 250 in attendance) and teams playing with very few fans (less than 250) in the stadium.
